---
title: "Tan.Query a Database with SQL"
author: "Jiaxing Tan"
date: '2023-02-10'
output: html_document
---

```{r}
library(pacman)
p_load(RSQLite)

dbfile = "MediaDB.db"

conn <- dbConnect(RSQLite::SQLite(),dbfile)
```

1. (5 pts) What are the last name, city, state, and country of all customers who made at least one purchase and live either in Brazil or Canada? Do not list duplicates.
```{sql connection=conn}
WITH t1 AS (
  SELECT lastname, city, state, country, customerid
  FROM customers 
  WHERE Country in ("Brazil", "Canada")
)
,t2 AS (
  SELECT customerid
  FROM invoices
  GROUP BY CustomerId
)

SELECT lastname, city, state, country
FROM t1
JOIN t2
ON t1.customerid = t2.customerid
```

2. (5 pts) What are the titles of each album and the number of tracks on each album? 
```{sql connection=conn}
SELECT t1.AlbumId, max(t1.Title) as title, count(distinct t2.TrackId) as num_tracks
FROM albums t1
LEFT JOIN tracks t2
ON t1.AlbumId = t2.AlbumId
GROUP BY t1.AlbumId
ORDER BY num_tracks DESC
```

3. (5 pts) List the number of tracks by genre, ordered from most to least, restricted to those genres for which there are at least five tracks.
```{sql connection=conn}
SELECT t2.GenreId, t2.Name as genre_name, count(*) as num_tracks
FROM tracks t1
JOIN genres t2
ON t1.GenreId = t2.GenreId
GROUP BY t2.GenreId, t2.Name
HAVING count(*) >= 5
ORDER BY num_tracks DESC
```


4. (10 pts) What is the average "tenure" of the employees in months, i.e., the time they have been working at our store? Use SQL to round the result to zero digit of precision, e.g., 23 instead of 23.4. Hint: SQLite does not support DATEDIFF but does support STRFTIME (see tutorial Links to an external site.). You need to only take year and month into account.
```{sql connection=conn}
WITH t1 AS (
  SELECT 
    strftime('%d', HireDate) AS Day,
    strftime('%m', HireDate) AS Month,
    strftime('%Y', HireDate) AS Year
  FROM employees
)

SELECT round(AVG( 
            (strftime('%Y', DATE('now')) - Year) * 12 + ( strftime('%m', DATE('now')) - Month)
            ), 0) as avg_tenure_months
FROM t1
```

5. (10 pts) What are the total number of unique customers for each Brazilian state (e.g., MG, SP, etc.), ordered alphabetically by state, who made at least one purchase (i.e., do not consider customers who never made a purchase)?
```{sql connection=conn}
WITH t1 AS (
  SELECT lastname, city, state, country, customerid
  FROM customers 
  WHERE Country = "Brazil"
)
,t2 AS (
  SELECT customerid
  FROM invoices
  GROUP BY CustomerId
)

SELECT state, count(distinct t1.customerid) as num_cust
FROM t1 
JOIN t2
ON t1.customerid = t2.customerid
GROUP BY state
ORDER BY state ASC

```



6. (10 pts) How many customers never made a purchase, i.e., they appear in the customers table but not in the invoices table.
```{sql connection=conn}
WITH t1 AS (
  SELECT lastname, city, state, country, customerid
  FROM customers 
)
,t2 AS (
  SELECT customerid
  FROM invoices
  GROUP BY CustomerId
)


SELECT count(distinct t1.customerid) as num_no_purchase_cust
FROM t1 
LEFT JOIN t2
ON t1.customerid = t2.customerid
WHERE t2.customerid IS NULL
```

7. (10 pts) How many albums contain the substring "symphony" in the album title?
```{sql connection=conn}
SELECT count(1) as ans
FROM albums
WHERE Title like '%symphony%'
```

8. (15 pts) What is the total "time" of all tracks per artist in hours, i.e., add up the "Milliseconds" column in tracks? Display the artist's name and the time in hours rounded to two decimals. Only consider artists who published at least one track.
```{sql connection=conn}
WITH t1 AS (
  SELECT t1a.AlbumId, t1b.Name, t1a.ArtistId
  FROM albums t1a
  JOIN artists t1b
  ON t1a.ArtistId = t1b.ArtistId
)
,t2 AS (
  SELECT ArtistId, t1.Name, t2a.Milliseconds
  FROM t1
  JOIN tracks t2a
  ON t1.AlbumId = t2a.AlbumId
)

SELECT ArtistId, name as artist_name, round(sum(Milliseconds * 1.0) / 3600000, 2) as totalhrs
FROM t2
GROUP BY ArtistId, Name
```

9. (15 pts) Which playlists are the longest (greater than 0), regardless of type and format? List the playlist name(s) and the time in two columns: one for hours and one for minutes.
```{sql connection=conn}
WITH t1 AS (
  SELECT t1a.TrackId, t1a.Milliseconds, t1b.PlaylistId
  FROM tracks t1a
  JOIN playlist_track t1b
  ON t1a.TrackId = t1b.TrackId
)
,t2 AS (
  SELECT trackid, milliseconds, t2a.PlaylistId, t2a.Name
  FROM t1
  JOIN playlists t2a
  ON t1.playlistid = t2a.PlaylistId
)
,t3 AS (
  SELECT PlaylistId, name, floor(sum(Milliseconds * 1.0) / 3600000) as totaltime_hr, sum(Milliseconds * 1.0) / 3600000 as totaltime 
  FROM t2
  GROUP BY PlaylistId, name
)

SELECT PlaylistId, name as playlist_name, totaltime_hr, floor((totaltime - totaltime_hr) * 60) as totaltime_min
FROM t3
WHERE totaltime = (SELECT max(totaltime) FROM t3)
```


10. (15 pts) Which artists have composed tracks in more than three different genres? List the artists name and the number of distinct genres they have compositions in. Do not list duplicates.
```{sql connection=conn}
WITH t1 AS (
  SELECT t1a.AlbumId, t1b.Name as artist_name, t1a.ArtistId
  FROM albums t1a
  JOIN artists t1b
  ON t1a.ArtistId = t1b.ArtistId
)
,t2 AS (
  SELECT ArtistId, t1.artist_name, t2a.TrackId, t2a.GenreId
  FROM t1
  JOIN tracks t2a
  ON t1.AlbumId = t2a.AlbumId
)

SELECT ArtistId, artist_name, count(distinct genreid) as gener_tally
FROM t2
GROUP BY ArtistId, artist_name
HAVING count(distinct genreid) >= 3
```



```{r}
dbDisconnect(conn)
```
