---
title: "Practicum I CS5200"
author: "Jiaxing Tan"
date: 'Spring 2023'
output: html_document
---

## Connect to the AWS RDS Database
```{r loadAllrequirements, warning=FALSE}
# install.packages("pacman")
library(pacman)
p_load(RMySQL, quietly=T)
p_load(tidyverse)
p_load(lubridate)
```

```{r connectToAWSRDSMySQLDatabase}
db_cred <- yaml::read_yaml('dbconfig.yml')
# Obviously not a good practice to exposure my credential, only for pedagogical purpose
db_user <- db_cred$db_user
db_password <- db_cred$db_password
db_name <- db_cred$db_name
db_host <- db_cred$db_host
db_port <- db_cred$db_port

mySQLConn <-  dbConnect(RMySQL::MySQL(), 
                      user = db_user, 
                      password = db_password,
                      dbname = db_name, 
                      host = db_host, 
                      port = db_port)
```

## Create Database

A. (5 pts / 0.5 hrs) Create a table incidents that stores wildlife strike incidents with this schema:
```{sql DropIncidents, connection=mySQLConn}
DROP TABLE IF EXISTS incidents
```

Because this is the first table, I will add foreign key constrains later.
```{sql CreateIncidents, connection=mySQLConn}
CREATE TABLE IF NOT EXISTS incidents (
    rid INTEGER NOT NULL AUTO_INCREMENT,
    `dep.date` DATE, 
    origin INTEGER, 
    airline INTEGER,
    aircraft VARCHAR(64), 
    `flight.phase` VARCHAR(32),
    altitude INTEGER, 
    conditions VARCHAR(64),
    warned BOOLEAN,
    PRIMARY KEY (rid),
    CONSTRAINT CHK_incident CHECK (`flight.phase` in ("takeoff", "landing", "inflight", "unknown") AND altitude > 0)
)
```

B. (5 pts / 0.5 hrs) Create a table that stores airports and states called airports and that follows this schema:
```{sql DropAirports,connection=mySQLConn}
DROP TABLE IF EXISTS airports
```

```{sql CreateAirports,connection=mySQLConn}
CREATE TABLE IF NOT EXISTS airports (
    aid INTEGER NOT NULL AUTO_INCREMENT,
    airportName VARCHAR(64) NOT NULL,
    airportCode VARCHAR(16),
    state VARCHAR(64) NOT NULL,
    PRIMARY KEY (aid)
)
```

C. (5 pts / 0.3 hrs) Link the incidents and airports tables via the origin foreign key in incidents to the primary key aid in airports.
```{sql CreateFKOriginOnIncidents, connection=mySQLConn}
ALTER TABLE incidents
ADD FOREIGN KEY (origin) REFERENCES airports(aid)
```

D. (5 pts / 0.5 hrs) Create a lookup table called conditions
```{sql DropConditions, connection=mySQLConn}
DROP TABLE IF EXISTS conditions
```

```{sql CreateConditions, connection=mySQLConn}
CREATE TABLE IF NOT EXISTS conditions (
    cid INTEGER NOT NULL AUTO_INCREMENT,
    `condition` VARCHAR(64) UNIQUE NOT NULL,
    explanation VARCHAR(128),
    PRIMARY KEY (cid)
)
```

Change the data type of conditions in table "incidents"
```{sql ChangeConditionsFromTextToInteger, connection=mySQLConn}
ALTER TABLE incidents
MODIFY conditions INTEGER
```

Link this lookup table to the incidents table through the conditions foreign key. 
```{sql CreateFKConditionsOnIncidents, connection=mySQLConn}
ALTER TABLE incidents
ADD FOREIGN KEY (conditions) REFERENCES conditions(cid)
```

E. (5 pts / 0.5 hrs) Create a table that stores airlines called airlines and that follows this schema.
```{sql DropAirlines, connection=mySQLConn}
DROP TABLE IF EXISTS airlines
```

```{sql CreateAirlines, connection=mySQLConn}
CREATE TABLE IF NOT EXISTS airlines (
    eid INTEGER NOT NULL AUTO_INCREMENT,
    airlineName VARCHAR(32) NOT NULL,
    airlineCode VARCHAR(8),
    flag VARCHAR(16),
    PRIMARY KEY (eid)
)
```

F. (3 pts / 0.3 hrs) Link the incidents and airlines tables via the airline foreign key in incidents to the primary key eid in airlines.
```{sql CreateFKAirlineOnIncidents, connection=mySQLConn}
ALTER TABLE incidents
ADD FOREIGN KEY (airline) REFERENCES airlines(eid)
```

G. Test
```{sql IncidentsFKConstrains, connection=mySQLConn, eval = FALSE}
SELECT
  TABLE_NAME,COLUMN_NAME,CONSTRAINT_NAME, REFERENCED_TABLE_NAME,REFERENCED_COLUMN_NAME
FROM
  INFORMATION_SCHEMA.KEY_COLUMN_USAGE
WHERE TABLE_NAME = 'incidents'
```


## Check the BirdStrikesData
```{r Check the structure of orginal dataset, warning=FALSE}
df_raw <- read.csv(file = "BirdStrikesData-V2.csv",
                   header = T,
                   stringsAsFactors = F,
                   nrows = 1000)

```

## Populate database with source data file 

1. Preprocess the raw dataframe:
  a. select required columns
  b. convert all NULL value, empty value into "unknown" or "1/1/1990 0:00" or "Airplane" or "Take-off run" or "N"
  c. convert data type of altitude for char to numeric.
  d. transfer flight phase to what we wanted
  e. convert date to Date data type
  f. convert warned data type to Boolean.
```{r preprocessing, warning=FALSE}
df_refine <- df_raw %>% 
  rename(., rid = ï..rid) %>%
  rename(., dep.date = flight_date) %>% 
  rename(., flight.phase = flight_phase) %>% 
  rename(., altitude = altitude_ft) %>% 
  rename(., conditions = sky_conditions) %>% 
  rename(., warned = pilot_warned_flag) %>% 
  select(., rid, dep.date, airline, aircraft, airport, flight.phase, origin, altitude, conditions, warned) %>% 
  as_tibble(.) %>% 
  mutate_at(., c("airline", "airport", "origin"), ~na_if(., '')) %>% 
  mutate_at(., c("airline", "airport", "origin"), ~replace_na(., "UNKNOWN")) %>% 
  mutate_at(., c("dep.date"), ~na_if(., '')) %>% 
  mutate_at(., c("dep.date"), ~replace_na(., "1/1/1990 0:00")) %>% 
  mutate_at(., c("aircraft"), ~na_if(., '')) %>% 
  mutate_at(., c("aircraft"), ~replace_na(., "Airplane")) %>%
  mutate_at(., c("flight.phase"), ~na_if(., '')) %>% 
  mutate_at(., c("flight.phase"), ~replace_na(., "Take-off run")) %>% 
  mutate_at(., c("warned"), ~na_if(., '')) %>% 
  mutate_at(., c("warned"), ~replace_na(., "N")) %>% 
  mutate_at(., c("altitude"), str_replace_all, 
            pattern = ",", 
            replacement = "") %>% 
  mutate_at(., c("altitude"), as.numeric) %>% 
  mutate(., flight.phase = case_when(
                                  flight.phase == "Take-off run" ~ "takeoff",
                                  flight.phase == "Landing Roll" | flight.phase == "Approach" ~ "landing",
                                  flight.phase == "Climb" | flight.phase == "Descent" ~ "inflight",
                                  TRUE ~ "unknown"
                                  )) %>% 
  mutate_at(., "dep.date", fast_strptime, format = "%m/%d/%Y %H:%M", lt = FALSE) %>% 
  mutate(., warned = case_when(
                              warned == "Y" ~ 1,
                              warned == "N" ~ 0,
                              TRUE ~ 0))  # RMySQL has a bug that take all logical to 0
```


2. Populate table "airlines"
```{r CreateLocalAirlines, warning=FALSE}
airlinesLC <- df_refine %>% group_by(., airline) %>% 
  summarise(.) %>% 
  rownames_to_column(.) %>% 
  mutate(airlinecode = NA) %>% 
  mutate(flag = NA) %>% 
  rename(., airlinename = airline) %>% 
  rename(., eid = rowname)
```

```{r WriteAirlines, warning=FALSE}
dbWriteTable(mySQLConn, "airlines", airlinesLC, overwrite = FALSE, 
             append = TRUE, row.names = FALSE)
```


3. Populate table "conditions"
```{r CreateLocalConditions, warning=FALSE}
conditions <- df_refine %>% group_by(., conditions) %>% 
  summarise(.) %>% 
  rownames_to_column(.) %>% 
  mutate(explanation = NA) %>% 
  rename(., cid = rowname) %>% 
  rename(., condition = conditions)
```

```{r WriteConditions, warning=FALSE}
dbWriteTable(mySQLConn, "conditions", conditions, overwrite = FALSE, 
             append = TRUE, row.names = FALSE)
```


4. Populate table "airports"
```{r CreateLocalAirports, warning=FALSE}
airports <- df_refine %>% group_by(., airport, origin) %>% 
  summarise(.) %>% 
  rownames_to_column(.) %>% 
  mutate(airportcode = NA, .after = airport) %>% 
  rename(., airportname = airport) %>% 
  rename(., state = origin) %>% 
  rename(., aid = rowname)
```

```{r WriteAirports, warning=FALSE}
dbWriteTable(mySQLConn, "airports", airports, overwrite = FALSE, 
             append = TRUE, row.names = FALSE)
```


5. Populate table "incidents"
Rename several columns for the convenience of later join strategy.
```{r CreateLocalIncidentsAndRenaming, warning=FALSE}
incidents <- df_refine %>% 
  rename(state = origin) %>% 
  rename(airline_src = airline) %>% 
  rename(conditions_src = conditions)
```

Match aid from table 'airports' into a new column origin.
```{r CreateColumnOrigin, warning=FALSE}
incidents <- incidents %>% 
  left_join(., airports, by = c("airport" = "airportname", "state"), keep = FALSE) %>% 
  select(., -airportcode) %>% 
  rename(origin = aid)
```

Match eid from table 'airlines' into a new column airline.
```{r CreateColumnAirline, warning=FALSE}
incidents <- incidents %>% 
  left_join(., airlinesLC, by = c("airline_src" = "airlinename"), keep = FALSE) %>% 
  select(., -airlinecode, -flag) %>% 
  rename(airline = eid)
```

Match cid from table 'conditions' into a new column 'conditions'
```{r CreateColumnConditions, warning=FALSE}
incidents <- incidents %>% 
  left_join(., conditions, by = c("conditions_src" = "condition"), keep = FALSE) %>% 
  select(., -explanation) %>% 
  rename(conditions = cid)
```

Select required columns in the right order
```{r FinalShuffleIncidents, warning=FALSE}
incidents <- incidents %>% select(rid, dep.date, origin, airline, aircraft, flight.phase, altitude, conditions, warned)
```

Write into MySQL table
```{r WriteIncidents, warning=FALSE}
dbWriteTable(mySQLConn, "incidents", incidents, overwrite = FALSE, 
             append = TRUE, row.names = FALSE)
```

## Show parts of each table
1. airlines
```{sql ShowAirlines, connection=mySQLConn}
SELECT *
FROM `airlines`
ORDER BY rand()
LIMIT 5
```

2. airports
```{sql ShowAirports, connection=mySQLConn}
SELECT *
FROM `airports`
ORDER BY rand()
LIMIT 5
```

3. conditions
```{sql ShowAirports, connection=mySQLConn}
SELECT *
FROM `airports`
ORDER BY rand()
LIMIT 5
```

4. incidents
```{sql Showincidents, connection=mySQLConn}
SELECT *
FROM `incidents`
ORDER BY rand()
LIMIT 5
```

## Create several SQLs against MySQL database



## Disconnect from Database
```{r disconnectedFromRDS}
dbDisconnect(mySQLConn)
```

