---
title: "Practicum I CS5200"
author: "Jiaxing Tan, Haoping Lin"
date: 'Spring 2023'
output: html_document
---

## Connect to the AWS RDS Database
```{r loadAllrequirements, warning=FALSE}
# install.packages("pacman")
library(pacman)
p_load(RMySQL, quietly=T)
p_load(tidyverse)
p_load(lubridate)
```

```{r connectToAWSRDSMySQLDatabase}
db_cred <- yaml::read_yaml('dbconfig.yml')
# Obviously not a good practice to exposure my credential, only for pedagogical purpose
db_user <- db_cred$db_user
db_password <- db_cred$db_password
db_name <- db_cred$db_name
db_host <- db_cred$db_host
db_port <- db_cred$db_port

mySQLConn <-  dbConnect(RMySQL::MySQL(), 
                      user = db_user, 
                      password = db_password,
                      dbname = db_name, 
                      host = db_host, 
                      port = db_port)
```

## Create Database

A. (5 pts / 0.5 hrs) Create a table incidents that stores wildlife strike incidents with this schema:
```{sql DropIncidents, connection=mySQLConn}
DROP TABLE IF EXISTS incidents
```

Because this is the first table, I will add foreign key constrains later.
```{sql CreateIncidents, connection=mySQLConn}
CREATE TABLE IF NOT EXISTS incidents (
    rid INTEGER NOT NULL AUTO_INCREMENT,
    `dep.date` DATE, 
    origin INTEGER, 
    airline INTEGER,
    aircraft VARCHAR(64), 
    `flight.phase` VARCHAR(32),
    altitude INTEGER, 
    conditions VARCHAR(64),
    warned BOOLEAN,
    PRIMARY KEY (rid),
    CONSTRAINT CHK_incident CHECK (`flight.phase` in ("takeoff", "landing", "inflight", "unknown") AND altitude >= 0)
)
```

B. (5 pts / 0.5 hrs) Create a table that stores airports and states called airports and that follows this schema:
```{sql DropAirports,connection=mySQLConn}
DROP TABLE IF EXISTS airports
```

```{sql CreateAirports,connection=mySQLConn}
CREATE TABLE IF NOT EXISTS airports (
    aid INTEGER NOT NULL AUTO_INCREMENT,
    airportName VARCHAR(64) NOT NULL,
    airportCode VARCHAR(16),
    state VARCHAR(64) NOT NULL,
    PRIMARY KEY (aid)
)
```

C. (5 pts / 0.3 hrs) Link the incidents and airports tables via the origin foreign key in incidents to the primary key aid in airports.
```{sql CreateFKOriginOnIncidents, connection=mySQLConn}
ALTER TABLE incidents
ADD FOREIGN KEY (origin) REFERENCES airports(aid)
```

D. (5 pts / 0.5 hrs) Create a lookup table called conditions
```{sql DropConditions, connection=mySQLConn}
DROP TABLE IF EXISTS conditions
```

```{sql CreateConditions, connection=mySQLConn}
CREATE TABLE IF NOT EXISTS conditions (
    cid INTEGER NOT NULL AUTO_INCREMENT,
    `condition` VARCHAR(64) UNIQUE NOT NULL,
    explanation VARCHAR(128),
    PRIMARY KEY (cid)
)
```

Change the data type of conditions in table "incidents"
```{sql ChangeConditionsFromTextToInteger, connection=mySQLConn}
ALTER TABLE incidents
MODIFY conditions INTEGER
```

Link this lookup table to the incidents table through the conditions foreign key. 
```{sql CreateFKConditionsOnIncidents, connection=mySQLConn}
ALTER TABLE incidents
ADD FOREIGN KEY (conditions) REFERENCES conditions(cid)
```

E. (5 pts / 0.5 hrs) Create a table that stores airlines called airlines and that follows this schema.
```{sql DropAirlines, connection=mySQLConn}
DROP TABLE IF EXISTS airlines
```

```{sql CreateAirlines, connection=mySQLConn}
CREATE TABLE IF NOT EXISTS airlines (
    eid INTEGER NOT NULL AUTO_INCREMENT,
    airlineName VARCHAR(32) NOT NULL,
    airlineCode VARCHAR(8),
    flag VARCHAR(16),
    PRIMARY KEY (eid)
)
```

F. (3 pts / 0.3 hrs) Link the incidents and airlines tables via the airline foreign key in incidents to the primary key eid in airlines.
```{sql CreateFKAirlineOnIncidents, connection=mySQLConn}
ALTER TABLE incidents
ADD FOREIGN KEY (airline) REFERENCES airlines(eid)
```

G. Test
```{sql IncidentsFKConstrains, connection=mySQLConn, eval = FALSE}
SELECT
  TABLE_NAME,COLUMN_NAME,CONSTRAINT_NAME, REFERENCED_TABLE_NAME,REFERENCED_COLUMN_NAME
FROM
  INFORMATION_SCHEMA.KEY_COLUMN_USAGE
WHERE TABLE_NAME = 'incidents'
```


## Check the BirdStrikesData
```{r Check the structure of orginal dataset, warning=FALSE}
df_raw <- read.csv(file = "BirdStrikesData-V2.csv",
                   header = T,
                   stringsAsFactors = F)

```

## Populate database with source data file 

1. Preprocess the raw dataframe:
  a. select required columns
  b. convert all NULL value, empty value into "unknown" or "1/1/1990 0:00" or "Airplane" or "Take-off run" or "N"
  c. convert data type of altitude for char to numeric.
  d. transfer flight phase to what we wanted
  e. convert date to Date data type
  f. convert warned data type to Boolean.
```{r preprocessing, warning=FALSE}
df_refine <- df_raw %>% 
  rename(., rid = ï..rid) %>%
  rename(., dep.date = flight_date) %>% 
  rename(., flight.phase = flight_phase) %>% 
  rename(., altitude = altitude_ft) %>% 
  rename(., conditions = sky_conditions) %>% 
  rename(., warned = pilot_warned_flag) %>% 
  select(., rid, dep.date, airline, aircraft, airport, flight.phase, origin, altitude, conditions, warned) %>% 
  as_tibble(.) %>% 
  mutate_at(., c("airline", "airport", "origin"), ~na_if(., '')) %>% 
  mutate_at(., c("airline", "airport", "origin"), ~replace_na(., "UNKNOWN")) %>% 
  mutate_at(., c("dep.date"), ~na_if(., '')) %>% 
  mutate_at(., c("dep.date"), ~replace_na(., "1/1/1990 0:00")) %>% 
  mutate_at(., c("aircraft"), ~na_if(., '')) %>% 
  mutate_at(., c("aircraft"), ~replace_na(., "Airplane")) %>%
  mutate_at(., c("flight.phase"), ~na_if(., '')) %>% 
  mutate_at(., c("flight.phase"), ~replace_na(., "Take-off run")) %>% 
  mutate_at(., c("warned"), ~na_if(., '')) %>% 
  mutate_at(., c("warned"), ~replace_na(., "N")) %>% 
  mutate_at(., c("altitude"), str_replace_all, 
            pattern = ",", 
            replacement = "") %>% 
  mutate_at(., c("altitude"), as.numeric) %>% 
  mutate(., flight.phase = case_when(
                                  flight.phase == "Take-off run" ~ "takeoff",
                                  flight.phase == "Landing Roll" | flight.phase == "Approach" ~ "landing",
                                  flight.phase == "Climb" | flight.phase == "Descent" ~ "inflight",
                                  TRUE ~ "unknown"
                                  )) %>% 
  mutate_at(., "dep.date", fast_strptime, format = "%m/%d/%Y %H:%M", lt = FALSE) %>% 
  mutate(., warned = case_when(
                              warned == "Y" ~ 1,
                              warned == "N" ~ 0,
                              TRUE ~ 0))  # RMySQL has a bug that take all logical to 0
```


2. Populate table "airlines"
```{r CreateLocalAirlines, warning=FALSE}
airlinesLC <- df_refine %>% group_by(., airline) %>% 
  summarise(.) %>% 
  rownames_to_column(.) %>% 
  mutate(airlinecode = NA) %>% 
  mutate(flag = NA) %>% 
  rename(., airlinename = airline) %>% 
  rename(., eid = rowname)
```

```{r WriteAirlines, warning=FALSE}
dbWriteTable(mySQLConn, "airlines", airlinesLC, overwrite = FALSE, 
             append = TRUE, row.names = FALSE)
```


3. Populate table "conditions"
```{r CreateLocalConditions, warning=FALSE}
conditions <- df_refine %>% group_by(., conditions) %>% 
  summarise(.) %>% 
  rownames_to_column(.) %>% 
  mutate(explanation = NA) %>% 
  rename(., cid = rowname) %>% 
  rename(., condition = conditions)
```

```{r WriteConditions, warning=FALSE}
dbWriteTable(mySQLConn, "conditions", conditions, overwrite = FALSE, 
             append = TRUE, row.names = FALSE)
```


4. Populate table "airports"
```{r CreateLocalAirports, warning=FALSE}
airports <- df_refine %>% group_by(., airport, origin) %>% 
  summarise(.) %>% 
  rownames_to_column(.) %>% 
  mutate(airportcode = NA, .after = airport) %>% 
  rename(., airportname = airport) %>% 
  rename(., state = origin) %>% 
  rename(., aid = rowname)
```

```{r WriteAirports, warning=FALSE}
dbWriteTable(mySQLConn, "airports", airports, overwrite = FALSE, 
             append = TRUE, row.names = FALSE)
```


5. Populate table "incidents"
Rename several columns for the convenience of later join strategy.
```{r CreateLocalIncidentsAndRenaming, warning=FALSE}
incidents <- df_refine %>% 
  rename(state = origin) %>% 
  rename(airline_src = airline) %>% 
  rename(conditions_src = conditions)
```

Match aid from table 'airports' into a new column origin.
```{r CreateColumnOrigin, warning=FALSE}
incidents <- incidents %>% 
  left_join(., airports, by = c("airport" = "airportname", "state"), keep = FALSE) %>% 
  select(., -airportcode) %>% 
  rename(origin = aid)
```

Match eid from table 'airlines' into a new column airline.
```{r CreateColumnAirline, warning=FALSE}
incidents <- incidents %>% 
  left_join(., airlinesLC, by = c("airline_src" = "airlinename"), keep = FALSE) %>% 
  select(., -airlinecode, -flag) %>% 
  rename(airline = eid)
```

Match cid from table 'conditions' into a new column 'conditions'
```{r CreateColumnConditions, warning=FALSE}
incidents <- incidents %>% 
  left_join(., conditions, by = c("conditions_src" = "condition"), keep = FALSE) %>% 
  select(., -explanation) %>% 
  rename(conditions = cid)
```

Select required columns in the right order
```{r FinalShuffleIncidents, warning=FALSE}
incidents <- incidents %>% 
  select(rid, dep.date, origin, airline, aircraft, flight.phase, altitude, conditions, warned)
```

Write into MySQL table
```{r WriteIncidents, warning=FALSE}
dbWriteTable(mySQLConn, "incidents", incidents, overwrite = FALSE, 
             append = TRUE, row.names = FALSE)
```


## Show parts of each table
1. airlines
```{sql ShowAirlines, connection=mySQLConn}
SELECT *
FROM `airlines`
ORDER BY rand()
LIMIT 5
```

2. airports
```{sql ShowAirports, connection=mySQLConn}
SELECT *
FROM `airports`
ORDER BY rand()
LIMIT 5
```

3. conditions
```{sql ShowAirports, connection=mySQLConn}
SELECT *
FROM `airports`
ORDER BY rand()
LIMIT 5
```

4. incidents
```{sql Showincidents, connection=mySQLConn}
SELECT *
FROM `incidents`
ORDER BY rand()
LIMIT 5
```

## Create several SQLs against MySQL database

#### Create a SQL query against your database to find the 10 states with the greatest number of incidents.
```{sql Q8, connection=mySQLConn}
SELECT a.state, count(*) as `number` 
FROM incidents i
JOIN airports a 
ON i.origin = a.aid
GROUP BY a.state
ORDER BY `number` DESC
LIMIT 10
```


#### Create a SQL query against your database to find the airlines that had an above average number bird strike incidents.
```{sql Q9, connection=mySQLConn}
WITH t1 AS (
  SELECT airline, count(1) as inci_tally
  FROM `incidents`
  GROUP BY airline
),
t2 AS (
  SELECT airline, inci_tally, AVG(inci_tally) over() as avg_inci_tally
  FROM t1
),
t3 AS (
  SELECT airline, inci_tally, avg_inci_tally
  FROM t2
  WHERE inci_tally > avg_inci_tally
)

SELECT t4.airlineName, t3.inci_tally, t3.avg_inci_tally
FROM t3
JOIN `airlines` t4
ON t3.airline = t4.eid
```

#### 10 Create a SQL query against your database to find the number of bird strike incidents by month and by flight phase (across all years).
```{r Q10, warning=FALSE}
sqlStatement <- "

select month(i.`dep.date`) as month, i.`flight.phase`, count(*) as number 
from incidents i
group by month(i.`dep.date`), i.`flight.phase`
order by month(i.`dep.date`)
"

bs = dbGetQuery(mySQLConn, sqlStatement)
head(bs, 6)
```

#### 11 Plot
```{r Q11, warning=FALSE}
bs.mod <- bs %>%  
  group_by(month) %>%  
  summarise(number = sum(number))
ggplot(bs.mod, aes(x=month, y=number)) +  
  geom_point() +  
  ggtitle("Number of Incidents by Months") +  
  xlab("Months") +  ylab("Num of Incidents") +  
  scale_x_continuous(limits = c(1, 12), breaks = c(1,2,3,4,5,6,7,8,9,10,11,12))
```


#### 12 Create a stored procedure in MySQL that adds a new incident to the database.
```{sql DropInsertIncidentsProcedure, connection=mySQLConn}
DROP PROCEDURE IF EXISTS insert_incident_sp
```

```{sql CreateInsertIncidentsProcedure, connection=mySQLConn}
CREATE PROCEDURE insert_incident_sp (
    IN `new_dep.date` DATE,
    -- IN new_origin INTEGER,
    IN new_airportname VARCHAR(64),
    IN new_state VARCHAR(64),
    -- IN airline INTEGER,
    IN new_airlinename VARCHAR(64),
    IN new_aircraft VARCHAR(64),
    IN `new_flight.phase` VARCHAR(32),
    IN new_altitude INTEGER,
    -- IN new_conditions INTEGER,
    IN new_conditionname VARCHAR(64),
    IN new_warned BOOLEAN
)
  BEGIN
    -- Check if origin exists in airports
    -- If exist, check state and airport name info
    -- Or insert a new row into table airports
    SELECT
  	aid
    INTO @aidflag
    FROM airports
    WHERE airportName = new_airportname;
  
    SELECT airportName
    INTO @pAirportName
    FROM airports
    WHERE airportName = new_airportname;
  
    SELECT state
    INTO @pState
    FROM airports
    WHERE airportName = new_airportname;
  
    -- Insert new record into airports or popup an error message
    IF (@aidflag is not null AND @pAirportName is not null AND @pState = new_state) THEN SELECT 1 WHERE FALSE; -- do nothing
    ELSEIF (@aidflag is not null AND @pAirportName is not null ) THEN SIGNAL sqlstate '02000'
              SET MESSAGE_TEXT = 'State values are not matched whereas airport name is found';
    ELSEIF @aidflag is null THEN
      INSERT INTO airports (airportName, state) VALUES (new_airportname, new_state);  -- Create new airport
      select aid into @aidflag from airports where airportName = new_airportname;  -- record new aid
    ELSE SIGNAL sqlstate '03000' SET MESSAGE_TEXT = 'Impossible case risen when checking airport foreign key';
    END IF;
  
    -- Check eid existence
    -- Check airline name info
    -- Insert new record if necessary
  
    SELECT
  	eid
    INTO @eidflag
    FROM airlines
    WHERE airlineName = new_airlinename;
  
    SELECT airlineName
    INTO @pAirlineName
    FROM airlines
    WHERE airlineName = new_airlinename;
  
      -- Insert new record into airports or popup an error message
    IF (@eidflag is not null AND @pAirlineName is not null) THEN SELECT 1 WHERE FALSE; -- do nothing
    ELSEIF @eidflag is null THEN
      INSERT INTO airlines (airlineName) VALUES (new_airlinename);  -- Create new airline
      select eid into @eidflag from airlines where airlineName = new_airlinename;  -- Record new eid
    ELSE SIGNAL sqlstate '03000' SET MESSAGE_TEXT = 'Impossible case risen when checking airlines foreign key';
    END IF;
  
    -- Check cid existence
    -- Check condition column info
    -- Insert new record if necessary
    SELECT
  	cid
    INTO @cidflag
    FROM conditions
    WHERE `condition` = new_conditionname;
  
    SELECT
  	`condition`
    INTO @condi
    FROM conditions
    WHERE `condition` = new_conditionname;
  
      -- Insert new record into airports or popup an error message
    IF (@cidflag is not null AND @condi is not null) THEN SELECT 1 WHERE FALSE; -- do nothing
    ELSEIF @cidflag is null THEN
      INSERT INTO conditions (`condition`) VALUES (new_conditionname);  -- Create new condition
      select cid into @cidflag from conditions where `condition` = new_conditionname;  -- Record new eid
    ELSE SIGNAL sqlstate '03000' SET MESSAGE_TEXT = 'Impossible case risen when checking condition foreign key';
    END IF;
  
    -- Insert new row into incidents table
    INSERT INTO incidents (`dep.date`, origin, airline, aircraft, `flight.phase`, altitude, conditions, warned)
        VALUES (`new_dep.date`, @aidflag, @eidflag, new_aircraft, `new_flight.phase`, new_altitude, @cidflag, new_warned);
  
  
    SET @aidflag = NULL;
    SET @pAirlineName = NULL;
    SET @pState = NULL;
    SET @eidflag = NULL;
    SET @pAirlineName = NULL;
    SET @cidflag = NULL;
    SET @condi = NULL;
  
  END
```

Test case
```{r NormalInsert, warning=FALSE}
sqlStatement = "
  CALL insert_incident_sp('2022-10-17', 'ADDISON', 'Texas'
    ,'UNITED AIRLINES', 'Airplane', 'landing', 200, 'No Cloud', 1)
"

rs = dbSendQuery(mySQLConn, sqlStatement)

# loop through any additional result sets and discard
while(dbMoreResults(mySQLConn) == TRUE) {
  dbNextResult(mySQLConn)    
}
```

```{r NewAirport, warning=FALSE}
sqlStatement = "
  CALL insert_incident_sp('2022-10-17', 'JAX', 'California'
    ,'UNITED AIRLINES', 'Airplane', 'landing', 200, 'No Cloud', 1)
"

rs = dbSendQuery(mySQLConn, sqlStatement)

# loop through any additional result sets and discard
while(dbMoreResults(mySQLConn) == TRUE) {
  dbNextResult(mySQLConn)
}
```

```{r exist_airport_but_incorrect_state, warning=FALSE}
# Error message expected
sqlStatement = "
  CALL insert_incident_sp('2022-10-17', 'ADDISON', 'California'
    ,'UNITED AIRLINES', 'Airplane', 'landing', 200, 'No Cloud', 1)
"

rs = tryCatch(
  {dbSendQuery(mySQLConn, sqlStatement)}
  ,error=function(cond) {
            message(cond)
            return(NA)
        })

# loop through any additional result sets and discard
while(dbMoreResults(mySQLConn) == TRUE) {
  dbNextResult(mySQLConn)
}
```

```{r NewAirline, warning=FALSE}
sqlStatement = "
  CALL insert_incident_sp('2022-10-17', 'ADDISON', 'Texas'
    ,'STARSHIP', 'Airplane', 'landing', 200, 'No Cloud', 1)
"
rs = dbSendQuery(mySQLConn, sqlStatement)

# loop through any additional result sets and discard
while(dbMoreResults(mySQLConn) == TRUE) {
  dbNextResult(mySQLConn)
}
```

```{r NewCondition, warning=FALSE}
sqlStatement = "
  CALL insert_incident_sp('2022-10-17', 'ADDISON', 'Texas'
      ,'UNITED AIRLINES', 'Airplane', 'landing', 200, 'Hurricane', 1)
"
rs = dbSendQuery(mySQLConn, sqlStatement)

# loop through any additional result sets and discard
while(dbMoreResults(mySQLConn) == TRUE) {
  dbNextResult(mySQLConn)
}
```

Note that there are new origin, new airline and new condition in the result.
```{sql CheckResult, connection=mySQLConn}
SELECT * FROM incidents where `dep.date` = '2022-10-17'
```

## Disconnect from Database
```{r disconnectedFromRDS}
dbDisconnect(mySQLConn)
```

