---
title: "Tan.CRDB.CS5200"
author: "Jiaxing Tan"
date: '2023-02-02'
output: html_document
---

```{r}
library(pacman)
p_load(RSQLite)

rootDB <- "databases"
dbfile = "ModuleDB.sqlitedb"

path <- getwd()
rootPath <- file.path(path, rootDB)
if (dir.exists(rootPath)) {
  message("Database root folder is already exist.")
  # return()
} else {
  if(!dir.create(rootPath)) stop("Given path in configDB is not exist. Please input a valid path for database")
}

conn <- dbConnect(RSQLite::SQLite(), file.path(rootPath, dbfile))
```

```{sql connection=conn}
PRAGMA foreign_keys = FALSE
```

```{sql connection=conn}
DROP TABLE IF EXISTS affiliation
```


```{sql connection=conn}
CREATE TABLE IF NOT EXISTS affiliation (
  aff_id integer primary key,
  aff_title text not null
)
```

```{sql connection=conn}
INSERT INTO affiliation (aff_id, aff_title)
VALUES 
    (1001, "PT"),
    (1002, "FT"),
    (2001, "Contract")

```


```{sql connection=conn}
DROP TABLE IF EXISTS author
```

```{sql connection=conn}
CREATE TABLE IF NOT EXISTS author (
  aid integer primary key,
  name text NOT NULL,
  email text NOT NULL,
  affiliation integer,
  FOREIGN KEY(affiliation) REFERENCES affiliation(aff_id)
)
```

```{sql connection=conn}
INSERT INTO author (aid, name, email, affiliation)
VALUES 
    (30, "Jon Snow", "jonsnow@example.com", 1002),
    (32, "Jaime Lannister", "jlannister@example.com", 1001),
    (33, "Tyrion Lannister", "tlannister@example.com", 2001),
    (40, "Arya Stark", "aryastark@example.com", 1002)

```

```{sql connection=conn}
SELECT * FROM author
```
```{sql connection=conn}
DROP TABLE IF EXISTS module;
```


```{sql connection=conn}
CREATE TABLE IF NOT EXISTS module (
  m_number text primary key,
  title text NOT NULL,
  length_in_min integer NOT NULL
)
```

```{sql connection=conn}
DROP TABLE IF EXISTS contribution;
```


```{sql connection=conn}

CREATE TABLE IF NOT EXISTS contribution (
  aid integer,
  m_number text,
  FOREIGN KEY(aid) REFERENCES author(aid),
  FOREIGN KEY(m_number) REFERENCES module(m_number)
)
```

Add triggers against module: 
1. Every time a new module row is inserted, automatically add a new record in contribution table with null in a_id attribute.
```{sql connection=conn}
CREATE TRIGGER contribution_after_module_insert 
   AFTER INSERT ON module
BEGIN
	INSERT INTO contribution (m_number)  -- ignore a_id
	VALUES (NEW.m_number)
	;
END
```

2. Every time a new module row is update, automatically modify all records in contribution table with the new m_number if that is what will be updated.
```{sql connection=conn}
CREATE TRIGGER contribution_after_module_update 
  AFTER UPDATE ON module
WHEN OLD.m_number <> NEW.m_number
BEGIN
  UPDATE contribution
  SET m_number = NEW.m_number
  WHERE m_number = OLD.m_number
	;
END
```

3. Every time a module row is deleted, automatically delete all records in contribution table with the m_number.
```{sql connection=conn}
CREATE TRIGGER contribution_after_module_delete
  AFTER DELETE ON module

BEGIN
  DELETE FROM contribution
  WHERE m_number = OLD.m_number
	;
END
```

4. Every time the last row of m_number in contribution table is deleted, automatically delete the record in module table with the m_number.
```{sql connection=conn}
CREATE TRIGGER module_after_contribution_delete
  AFTER DELETE ON contribution
WHEN OLD.m_number not in (
      SELECT m_number FROM contribution
      WHERE m_number = OLD.m_number
    )

BEGIN
  DELETE FROM module
  WHERE m_number = OLD.m_number
	;
END
```

Test:
```{sql connection=conn}
INSERT INTO module (m_number, title, length_in_min)
VALUES ('CS5200', 'Database', 1200)
```

```{sql connection=conn}
SELECT * FROM contribution
```
Then modified the aid
```{sql connection=conn}
UPDATE contribution
SET aid = 30
WHERE m_number = "CS5200"
```




```{sql connection=conn}
SELECT * FROM sqlite_schema
WHERE type='trigger'
ORDER BY name;
```

```{r}
dbDisconnect(conn)
```

